A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950

### 2022-02-07 paper4.Rmd
first commit

---
title: "A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950, including the Stensele station and the inland"
author: "Göran Broström and Tommy Bengtsson"
date: "`r Sys.time()`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  bookdown::word_document2:
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
citation_package: natbib
classoption: titlepage
bibliography: bio.bib
titlepage: yes
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{a4wide}
#- \usepackage[figuresonly]{endfloat}
#- \renewcommand{\efloatseparator}{\mbox{}}
abstract: "The effect of extreme temperatures on infant mortality in the Umeå region 1895-1950 is studied in a biometric analysis setting. More precisely, the effect of climate and weather, measured by temperature, on infant mortality is investigated. It turns out that climate is more important than weather, low average temperatures are more important than temporary dips in temperature. The investigated geographical region is situated too far north for bad effects of high temperatures to show."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path='figs/',fig.height=3, comment=NA)
knitr::opts_knit$set(global.par = TRUE)
##
library(knitr)
library(eha)
##par(las = 1, cex.axis = 0.7, cex.lab = 0.7)
g <- function(x) (log(1 + x))^3 # x in days.
dg <- function(x) (3 * (log(1 + x))^2) / (1 + x)
ginv <- function(y) exp(y^(1/3)) - 1
```

@@

# Introduction

@@

One important reason for studying neonatal and postneonatal mortality separately is the 
empirical findings by Bourgeois-Pichat [@bp51a;  @bp51b] about endogenous and exogenous 
mortality and the log-cube transform.

@@

Individual data with all births between 1 January 1895 and 31 December 1950 in two 
coastal areas of north Sweden, Skellefteå (51560 births) and Umeå (31213 births).
They were followed until death or age 365 days, whichever came first. 
The following *static* characteristics were observed on each child:

@@

```{r mortstat, fig.cap = "Crude infant mortality by week of year, Umeå/Skellefteå 1895--1950.", fig.height = 3}
infdat <- readRDS("data/infdat3.rds")
tid <- toTpch(Surv(enter, exit, event) ~ week, data = infdat, cuts = c(0, 1))
y <- tid$event / tid$exposure
par(las = 1, cex.axis = 0.7, cex.lab = 0.7)
plot(1:52, y, type = "b", col = "blue", axes = FALSE, ylim = c(0, max(y)),
     xlab = "Week", ylab = "IMR", cex = 0.7, lwd = 1.5)
axis(1, at = c(1, 10, 20, 30, 40, 52))
axis(2, las = 1, at = c(0, 0.03, 0.06, 0.09))
abline(h = c(0.05, 0.07, 0.09, 0.11), lty = 3)
abline(h = 0)
box()
```

@@

The analyses are performed on the *log-cube* scale, following the hints of  
Bourgeois-Pichat [@bp51a; @bp51b], with proportional hazards modelling. Note that
the property of proportional hazards are preserved under a strictly monotone increasing time transform.
However, the estimates of baseline distribution characteristics will change, of course.

@@

```{r prepare}
source("R/ltx2.R")
infdat$parity <- cut(infdat$parity, c(-2, 1.5, 4.5, 19), labels = c("1", "2-4", "5+"))
infdat$period <- cut(infdat$birthdate, c(1895, 1914, 1935, 1951), dig.lab = 5)
##levels(infdat$subreg) <- c("ume", "ske")
##infdat$season <- cut(infdat$week, c(-1, 9.5, 22.5, 35.5, 48.5, 52.5), labels = c("winter", "spring", "summer", "fall", "winter"))
##infdat$season <- as.factor(infdat$season)
## New:
infdat$summer <- infdat[infdat$week %in% 14:39, ]
infdat$emeantemp <- round(infdat$emeantemp)
levels(infdat$socBranch) <- c("office", "farming", "office", "worker", "worker") # NOTE!
```

@@

In accordance with the results of @bp51a and @bp51b, we separate the investigation into 
two parts, *neonatal* and *postneonatal* mortality.

@@

```{r introneo}
neo <- age.window(infdat, c(0, 28/ 365))
```

The analyses are split into two parts by season, *winter + spring* is one, and *summer + fall* the other.

@@

```{r neoanalw, results = 'asis'}
##library(survival)
winter <- neo[!neo$summer, ]
##winter$season <- factor(winter$season)
##neo$exit <- g(neo$exit * 365) # Skip the log-cube transformation
##neo$enter <- g(neo$enter * 365)
fit.cr <- coxreg(Surv(enter, exit, event) ~  highTemp + aver + emeantemp + 
                     sex + subreg + illeg + parity + socBranch,
                 data = winter)
dr <- drop1(fit.cr, test = "Chisq")
ltx2(fit.cr, dr = dr, caption = "Neonatal mortality, October to March. Adjusted for social branch, sex, illegitimacy, parity, time period, and subregion.", 
     label = "tab:neoanalw", keep = NULL)
```

@@

```{r neoanals, results = 'asis'}
##library(survival)
summer <- neo[neo$summer, ]
##summer$season <- factor(summer$season)
##neo$exit <- g(neo$exit * 365) # Skip the log-cube transformation
##neo$enter <- g(neo$enter * 365)
fit.cr <- coxreg(Surv(enter, exit, event) ~  aver + emeantemp + 
                     sex + subreg + illeg + parity + socBranch,
                 data = summer)
dr <- drop1(fit.cr, test = "Chisq")
ltx2(fit.cr, dr = dr, caption = "Neonatal mortality, April to September. Adjusted for social branch, sex, illegitimacy, parity, time period, and subregion.", 
     label = "tab:neoanals", keep = NULL)
```

@@

This is the simplest part, because @bp51a predicts that the baseline distribution (given the log-cube
transformation **g**) is *exponential*, that is, the hazard function is *constant*.
To give substance to this claim, a short introduction to the ideas of Bourgeois-Pichat is
given.

### Bourgeois-Pichat and the log-cube transform

The *log-cube transform g*, suggested by @bp51a, is defined as

\begin{equation}
g(t) = \log^3(1 + t), \quad 0 \le t \le 365,
\end{equation}

where $t$ is age in days. A graph of the transform is shown in Figure \@ref(fig:gplot).

```{r gplot, fig.cap = "The g transform.",fig.height=3.2}
x <- seq(0, 365, length = 1000)
y <- g(x)
plot(x, y, type = "l", ylab = "g(t)", xlab = "t (days)", col = "blue", axes = FALSE)
abline(h = 0, v = 0)
##
lines(c(28, 28), c(0, g(28)), lty = 3)
lines(c(0, 28), c(g(28), g(28)), lty = 3)
##
lines(c(183, 183), c(0, g(183)), lty = 3)
lines(c(0, 183), c(g(183), g(183)), lty = 3)
##
lines(c(365, 365), c(0, g(365)), lty = 3)
lines(c(0, 365), c(g(365), g(365)), lty = 3)
##
axis(1, at = c(0, 28, 183, 365))
ycor <-  c(0, g(28), g(183), g(365))
axis(2, at = ycor, labels = round(ycor), las = 1)
box()
```

It was used as a tool for dividing infant mortality 
into *endogenous* and *exogenous* causes, and a an important part in that venue was
the observations (i) postneonatal mortality is purely exogenous, and (ii) the distribution 
of exogenous mortality is *truncated uniform* (right truncated at $t = g(365)$) on the
$g$-transformed time scale.

Today, assumption (ii) is slightly outdated, for three reasons: (a) The uniform distribution
is not a very practical tool in survival analysis, (b) replacing the uniform with an
*exponential* not only fits better (usually), it also is a survival distribution that
is extremely easy to work with in survival analysis, and (c) "all models are wrong, but some are useful [@box76]".

```{r getpneo}
pneo <- age.window(infdat, c(28 / 365, 1))
gpneo <- pneo
gpneo$enter <- g(gpneo$enter * 365)
gpneo$exit <- g(gpneo$exit * 365)
##
oj <- system.time(gtpneo <- toTpch(Surv(enter, exit, event) ~ period + sex + socBranch + illeg + parity + 
                    lowTemp + highTemp + summer + aver
                + emeantemp + subreg, 
                data = gpneo, cuts = g(c(28, 365))))
```

For now, let us look at the basic distribution of postneonatal life on the $g$ time
scale. We do it nonparametrically by calculating the *Nelson-Aalen* estimator of the
cumulative hazards function. See Figure \@ref(fig:nap), where the two curves are indistinguishable.

```{r nap, fig.cap = "Nelson-Aalen estimator and exponential cumulative hazards of postneonatal survival on the g scale.",fig.height=3.2}
par(cex = 0.8)
fit <- coxreg(Surv(enter, exit, event) ~ 1, data = gpneo)
xx <- hazards(fit)
plot(xx[[1]][, 1], cumsum(xx[[1]][, 2]), type = "s", xlab = "Age (days)",
     col = "blue", ylab = "Cumulative hazards", axes = FALSE)
axis(1, at = g(c(28, 183, 365)), labels = c(28, 183, 365))
axis(2)
box()
fit.w <- tpchreg(oe(event, exposure) ~ 1, data = gtpneo, time = age)
zz <- hazards(fit.w)
lines(zz$x, zz$y, col = "red", lty = 2)
legend("topleft", legend = c("Nelson-Aalen", "Exponential"), lty = 1:2, col = c("blue", "red"))
```

The survival data set we work with here contains 102 thousand infants and 4.9 million records,
an average of 48 records per infant, one record per calendar week in life. In total there are 
`r sum(gpneo$event)` deaths, and a typical Cox regression execution time is slightly
below one minute. A "naive" application of an exponential regression model to the same data
is typically using around six minutes.

However, with the exponential model (or generally, the *piecewise constant hazard* (pch)) in mind, data can be heavily reduced by *tabulation*, and the
statistical *sufficiency principle* will guarantee that we get identical results.
The tabulation itself takes about 20 seconds, and the analysis after that about 15 seconds, a
total of 35 seconds. But the tabulation usually only needs to be done once. More about this later.

@@

```{r postanalw, results='asis'}
winter <- gtpneo[gtpneo$summer == 0, ]
##winter$season <- factor(winter$season)

fit.tpch <- tpchreg(oe(event, exposure) ~ lowTemp + highTemp + aver + emeantemp + sex + socBranch  + 
                        illeg + parity + subreg, 
                    data = winter, time = age)
dr <- drop1(fit.tpch, test = "Chisq")
ltx2(fit.tpch, dr = dr, caption = "Postneonatal mortality, October to March. Adjusted for social branch, sex, illegitimacy, parity, time period, and subregion.", label = "tab:postanalw", keep = NULL)
```

@@

```{r postanals, results='asis'}
summer <- gtpneo[gtpneo$summer == 1, ]
##winter$season <- factor(winter$season)

fit.tpch <- tpchreg(oe(event, exposure) ~ lowTemp + highTemp + aver + emeantemp + sex + socBranch  + 
                        illeg + parity + subreg, 
                    data = summer, time = age)
dr <- drop1(fit.tpch, test = "Chisq")
ltx2(fit.tpch, dr = dr, caption = "Postneonatal mortality, April to September. Adjusted for social branch, sex, illegitimacy, parity, time period, and subregion.", label = "tab:postanals", keep = NULL)
```

@@

# The biometric analysis of infant mortality

A modern version of the Bourgeois-Pichat biometric modelling [@bp51a; @bp51b] is given here,
*without* the "log-cube transform". A first try is fit a *Weibull* distribution to the time
to death for an infant alive at age 28 /365 years (28 days), that is, time is measured from
age 28 days until death or time reaches 365 - 28 days, when follow-up is censored.
In **R** language, with the package *eha* [@eha], we try

```
births <- readRDS("data/births.rds")
pbirths <- age.window(births, c(28 / 365, 1)
fit.w <- phreg(Surv(enter - 28/365, exit - 28/365, event) ~ 1, data = pbirths)
fit <- coxreg(Surv(enter - 28/365, exit - 28/365, event) ~ 1, data = pbirths)
```

where `phreg` performs the Weibull regression (no covariates) and `coxreg` the Cox regression. In 
Figure \@ref(fig:bpprim) the result is shown in the form of the two cumulative hazards estimates.

```{r bpprim, fig.cap = "Postneonatal survival, Weibull vs. Nelson-Aalen."}
births <- readRDS("data/births.rds")
pbirths <- age.window(births, c(28 / 365, 1))
fit.w <- phreg(Surv(enter - 28 / 365, exit - 28 / 365, event) ~ 1, data = pbirths)
fit <- coxreg(Surv(enter - 28 / 365, exit - 28 / 365, event) ~ 1, data = pbirths)
plot(fit, fn = "cum", col = "blue")
xx <- hazards(fit.w)
lines(xx$x, xx$y, col = "red", lty = 2)
```

Not a very good fit. However, we need a *left-truncated Weibull distribution*, that is, assuming that
*exogeneous infant mortality* is Weibull distributed.

```
fit.w <- phreg(Surv(enter, exit, event) ~ 1, data = pbirths)
fit <- coxreg(Surv(enter, exit, event) ~ 1, data = pbirths)
```
Note the difference: No subtraction of the 28 days from the time variables, which implies that the time origin
is *birth* and *not* the 28th day of life. Thus our data are *left truncated* at 28 days, and the 
Weibull attempt estimates neonatal exogeneous mortality through *extrapolation*. The Cox regression approach
is a non-parametric procedure, and thus not capable of any extrapolated estimation, so neonatal
mortality will simply be taken as zero.

```{r ltruw, fig.cap = "Postneonatal survival, truncated Weibull vs. Nelson-Aalen."}
fit.w <- phreg(Surv(enter, exit, event) ~ 1, data = pbirths)
fit <- coxreg(Surv(enter, exit, event) ~ 1, data = pbirths)
oldpar <- par(mfrow = c(1, 2))
##check.dist(fit, fit.w)
cc <- hazards(fit)
plot(cc[[1]][, 1], cumsum(cc[[1]][, 2]), type = "l", col = "blue", ylim = c(0, 0.07), axes = FALSE,
     xlab = "Age", ylab = "Cumulative hazards")
axis(1, at = c(28 / 365, 1), labels = c("28/365", "1"))
axis(2, las = 1)
box()
xx <- hazards(fit.w)
lines(xx$x, xx$y, col = "red", lty = 2)
plot(cc[[1]][, 1], cumsum(cc[[1]][, 2]), type = "l", col = "blue", ylim = c(0, 0.07), axes = FALSE,
     xlab = "Age", ylab = "Cumulative hazards")
##plot(fit, fn = "cum", col = "blue", ylim = c(0, 0.07), axes = FALSE)
axis(1, at = c(28 / 365, 1), labels = c("28/365", "1"))
axis(2, las = 1)
box()
lines(xx$x, xx$y - xx$y[1], col = "red", lty = 2)
par(oldpar)
```

 Look here (Figure \@ref(fig:ltruw))! A perfect fit, 
just like the one under the $g$- transform. But the Weibull curve
does not start at zero (left hand panel), however it is parallell 
to the Nelson-Aalen estimator, and the right hand panel shows what happens if we subtract 
the start value from the Weibull curve and redraw.

## The biometric analysis of  exogeneous and endogeneous infant mortality

So, finally, we show the full procedure how to perform the "biometric analysis" *without*
the log-cube transform. It simply consists of two steps:

1.    Run a Cox regression on the full data set.
2.    Run a *left truncated* (at 28 days) Weibull regression.
3.    Plot the two cumulative hazards functions in the same graph and read
      their crossings of the vertical line at age = 1 year (365 days). Figure \@ref(fig:crossings).

```{r crossings, fig.cap = "Biometric analysis.", fig.height = 4}
fit.c <- coxreg(Surv(enter, exit, event) ~ 1, data = births)
fit.w <- phreg(Surv(enter, exit, event) ~ 1, data = pbirths)
cc <- hazards(fit.c)
plot(c(0, cc[[1]][, 1]), c(0, cumsum(cc[[1]][, 2])), type = "l", col = "blue", axes = FALSE,
     xlab = "Age (days)", ylab = "Cumulative hazards", ylim = c(0, 0.08))
axis(1, at = c(0, 28 / 365, 1), labels = c(0, 28, 365))
axis(2, las = 1, at = round(c(0, max(cumsum(cc[[1]][, 2])), max(xx$y)), 3))
##abline(v = 1)
box()
xx <- hazards(fit.w)
lines(xx$x, xx$y, col = "red", lty = 2)
abline(h = 0, v = 0)
abline(v = 1, lty = 3)
abline(v = 28 / 365, lty = 3)
abline(h = c(max(cumsum(cc[[1]][, 2])), max(xx$y)), lty = 3)
```

So, total mortality is 0.076, exogeneous mortality is 0.063, and endogeneous mortality
is the difference, 0.013, or 17 per cent of the total mortality.

Now, try the original "biometric" approach and compare *(Exercise)*.

>>>>

---
title: "A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950"
author: "Göran Broström and Tommy Bengtsson"
date: "`r Sys.time()`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  bookdown::word_document2:
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
citation_package: natbib
classoption: titlepage
bibliography: bio.bib
titlepage: yes
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{a4wide}
#- \usepackage[figuresonly]{endfloat}
#- \renewcommand{\efloatseparator}{\mbox{}}
abstract: "The effect of extreme temperatures on infant mortality in the Umeå and Skellefteå regions 1895-1950 is studied in a biometric analysis setting. More precisely, the effect of climate and weather, measured by temperature, average and extremes, on infant mortality is investigated. It turns out that climate (average) is more important than weather (extremes), low average temperatures are more important than temporary dips in temperature."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path='figs/',fig.height=3, comment=NA)
knitr::opts_knit$set(global.par = TRUE)
##
library(knitr)
library(eha)
```

@@

```{r readdata}
births <- readRDS("data/births.rds")
infdat  <- readRDS("data/infdat4.rds")
```
# Introduction

@@

@@

Individual data with all births between 1 January 1895 and 31 December 1950 in two 
coastal and one inland areas of north Sweden, Skellefteå (51560 births) and Umeå (31213 births).
They were followed until death or age one year, whichever came first. 
The following *static* characteristics were observed on each child:

@@

```{r mortstat, fig.cap = "Crude infant mortality by week of year, Umeå/Skellefteå 1895--1950.", fig.height = 3}
##infdat <- readRDS("data/infdat4.rds") Already done
tid <- toTpch(Surv(enter, exit, event) ~ week, data = infdat, cuts = c(0, 1))
y <- tid$event / tid$exposure
par(las = 1, cex.axis = 0.7, cex.lab = 0.7)
plot(1:52, y, type = "b", col = "blue", axes = FALSE, ylim = c(0, max(y)),
     xlab = "Week", ylab = "IMR", cex = 0.7, lwd = 1.5)
axis(1, at = c(1, 10, 20, 30, 40, 52))
axis(2, las = 1, at = c(0, 0.03, 0.06, 0.09))
abline(h = c(0.05, 0.07, 0.09, 0.11), lty = 3)
abline(h = 0)
box()
```

@@

@@

```{r prepare}
source("R/ltx2.R")
infdat$parity <- cut(infdat$parity, c(-2, 1.5, 4.5, 19), labels = c("1", "2-4", "5+"))
##infdat$period <- cut(infdat$birthdate, c(1895, 1914, 1935, 1951), dig.lab = 5)
##levels(infdat$subreg) <- c("ume", "ske")
##infdat$season <- cut(infdat$week, c(-1, 9.5, 22.5, 35.5, 48.5, 52.5), labels = c("winter", "spring", "summer", "fall", "winter"))
##infdat$season <- as.factor(infdat$season)
## New:
infdat$summer <- infdat$week %in% 14:39
infdat$emeantemp <- round(infdat$emeantemp)
levels(infdat$socBranch) <- c("office", "farming", "office", "worker", "worker") # NOTE!
```

@@

We separate the investigation into two parts, *neonatal* and *postneonatal* mortality. But first,
a joint analysis.

```{r joint, cache = TRUE}
library(survival)
neo <- age.window(infdat, c(0, 28 / 365))
neo$ageIvl <- "neo"
pneo <- age.window(infdat, c(28/365, 1))
pneo$ageIvl <- "postneo"
all <- rbind(neo, pneo)
all$ageIvl <- factor(all$ageIvl)
##all$enter <- round(all$enter * 365)
##all$exit <- round(all$exit * 365)
fit.all <- coxph(Surv(enter, exit, event) ~ strata(ageIvl) * (aver + emeantemp + lowTemp + highTemp + socBranch + socStatus + sex + period),
                 data = all)
dr.all <- drop1(fit.all, test = "Chisq")

inftab <- toTpch(Surv(enter, exit, event) ~ subreg + sex + birthmonth + period + socBranch + socStatus + parity + 
                     aver + emeantemp + lowTemp + highTemp + summer + ageIvl, data = all,
                 cuts = c(0, 3, 7, 14, 28, 56, 84, 145, 207, 286, 365) / 365)
tabneo <- inftab[inftab$ageIvl == "neo", ]
tabpost <- inftab[inftab$ageIvl == "postneo", ]
tabpost$age <- factor(tabpost$age)
tabneo$age <- factor(tabneo$age)
##system.time(tabneo.fit <- tpchreg(oe(event, exposure) ~ aver + emeantemp + highTemp + lowTemp + socBranch + socStatus, 
  ##               data = tabneo, time = age))
##system.time(tabpost.fit <- tpchreg(oe(event, exposure) ~ aver + emeantemp + highTemp + lowTemp + socBranch + socStatus, 
  ##               data = tabpost, time = age))
```

```{r shownonprop}
tbl(dr.all, caption = "Proportionality tests, neonatal vs. postneonatal.", rownames = TRUE)
```

@@

The analyses are split into two parts by season, *winter* is one, and *summer* the other.
Summer covers the months April to September, and winter the rest of the months.

@@

```{r neoanalw, results = 'asis'}
##library(survival)
winter <- neo[!neo$summer, ]
##winter$season <- factor(winter$season)
##neo$exit <- g(neo$exit * 365) # Skip the log-cube transformation
##neo$enter <- g(neo$enter * 365)
fit.cr <- coxreg(Surv(enter, exit, event) ~  highTemp + aver + emeantemp + 
                     sex + subreg + illeg + parity + socBranch + period,
                 data = winter)
dr <- drop1(fit.cr, test = "Chisq")
ltx2(fit.cr, dr = dr, caption = "Neonatal mortality, October to March. Adjusted for social branch, sex, illegitimacy, parity, time period, and subregion.", 
     label = "tab:neoanalw", keep = NULL)
```

@@

```{r neoanals, results = 'asis'}
##library(survival)
#summer <- neo[neo$summer, ]
##summer$season <- factor(summer$season)
##neo$exit <- g(neo$exit * 365) # Skip the log-cube transformation
##neo$enter <- g(neo$enter * 365)
fit.cr <- tpchreg(oe(event, exposure) ~  aver + emeantemp + 
                     sex + subreg  + socBranch + period,
                 data = inftab[inftab$summer, ], time = age)
dr <- drop1(fit.cr, test = "Chisq")
ltx2(fit.cr, dr = dr, caption = "Neonatal mortality, April to September. Adjusted for social branch, sex, illegitimacy, parity, time period, and subregion.", 
     label = "tab:neoanals", keep = NULL)
```

@@

@@

```{r postanalw, results='asis'}
##tpneo <- toTpch(Surv(enter, exit, event) ~ lowTemp + highTemp + aver + emeantemp + sex + socBranch + illeg + parity
  ##              + subreg + period, data = infdat, cuts = c(0, 1))
##winter <- gtpneo[gtpneo$summer == 0, ]
##winter$season <- factor(winter$season)

fit.tpch <- tpchreg(oe(event, exposure) ~ lowTemp + highTemp + aver + emeantemp + sex + socBranch  + 
                        parity + subreg + period + summer, data = tabpost[!tabpost$summer, ], time = age)
dr <- drop1(fit.tpch, test = "Chisq")
ltx2(fit.tpch, dr = dr, caption = "Postneonatal mortality, October to March. Adjusted for social branch, sex, illegitimacy, parity, time period, and subregion.", label = "tab:postanalw", keep = NULL)
```

@@

```{r postanals, results='asis'}
##summer <- gtpneo[gtpneo$summer == 1, ]
##winter$season <- factor(winter$season)

fit.tpch <- tpchreg(oe(event, exposure) ~ lowTemp + highTemp + aver + emeantemp + sex + socBranch  + 
                        parity + subreg + period, 
                    data = tabpost[tabpost$summer, ], time = age)
dr <- drop1(fit.tpch, test = "Chisq")
ltx2(fit.tpch, dr = dr, caption = "Postneonatal mortality, April to September. Adjusted for social branch, sex, illegitimacy, parity, time period, and subregion.", label = "tab:postanals", keep = NULL)
```

@@
